%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../index"
%%% End:

% Recommended:
% $(t, w)$
% - Why can $t$-people reconstruct the key?
% - Why cannot less than $t$-people reconstruct the key?

\subsection*{Agenda}
\begin{enumerate}
\item The issue
\item Shamir's threshold scheme
\item Key re-construction
\end{enumerate}

\subsection{The issue}
Imagine you have encrypted some data using a
symmetric encryption algorithm. What if you then lost the secret key?
If the encryption is strong, your data is unrecoverable.

To avoid this you could have given the secret key to one or more
friends, as a backup. The problem is now is that your friends might
also lose the key. Also, each of your friends can decrypt your
important data.

What if you could give  only shares of the secret key to your
friends, but only need a subset of the shares to reconstruct the
secret key? Now none of your friends can alone decrypt you data, and
it doesn't matter if a friend or two loses their share. This is
\emph{secret sharing}.

\subsection{Shamir's threshold scheme}
\subsubsection*{Setup}
\begin{itemize}
\item $(t, w)$, where $t$ is the threshold and $w$ is the number of shareholders
\item Participants: Keyholder $H$, shareholders $S_1, \ldots, S_w$
\item Secret key: $k$
\item H chooses:
  \begin{itemize}
  \item a prime $p$, such that $p \ge w + 1$
  \item a random polynomial
    \[
    f(x) = k + \sum\limits_{j=1}^{t-1} c_jx^j \mod p
    \]
    of degree $t-1$, where $c_j$s are chosen at random
  \end{itemize}
\item $H$ computes $(x_i, f(x_i))$ for $i  = 1, \ldots, t$
\end{itemize}

\subsubsection*{Re-construction}
\begin{figure}[H]
  \begin{equation}\label{eq:lagrange}
    f(x) = \sum\limits_{i=1}^t y_1 \prod\limits_{j=1, j\not=i}^t \frac{x-x_j}{x_i-x_j}
  \end{equation}
  \label{asdf}
  \caption{Lagrange Interpolation Formula}
\end{figure}

\begin{itemize}
\item Input: $t$ shares $(x_i, y_i)$ for $i = 1, \ldots, t$
\item $H$ re-constructs the secret key $k$ by using Lagrange Interpolation Formula (\ref{eq:lagrange})
  \[ k = f(0) = \sum\limits_{i=1}^t y_1 \prod\limits_{j=1, j\not=i}^t \frac{x_j}{x_j-x_i} \]
\end{itemize}

\textbf{$t$ shareholders} \\
There is only one polynomial $f(x)$ over $\mathbb{Z}_p$ of degree at
most $t-1$, such that $f(x_i) = y_i$ for $i = 1,\ \dots,\
t$. Therefore, $t$ shareholders with the shares $\{(x_1,y_1),\
(x_2,y_2),\ \ldots,\ (x_t,y_t)\}$, can re-construct the polynomial
$f(x)$ using the Lagrange Interpolation Formula~(\ref{eq:lagrange}) and
compute the secret key for $f(0)$.

Further, if we describe the shares as follows
\begin{equation}
  \begin{pmatrix}
    1      & x_1    & x_1^2    & \cdots & x_1^{t-1} \\
    1      & x_2    & x_2^2    & \cdots & x_2^{t-1} \\
    \cdots & \cdots & \cdots  & \cdots & \cdots \\
    1      & x_{t-1} & x_{t-1}^2 & \cdots & x_{t-1}^{t-1} \\
    1      & x_t    & x_t^2    & \cdots & x_t^{t-1}
  \end{pmatrix}
  \begin{pmatrix}
    c_0 \\
    c_1 \\
    c_2 \\
    \cdots \\
    c_{t-1}
  \end{pmatrix} =
  \begin{pmatrix}
    y_0 \\
    y_1 \\
    y_2 \\
    \cdots \\
    y_t
  \end{pmatrix} \mod p
\end{equation}

Since $p$ is assumed to be a prime, and since $x_i \not= x_j$ for $1 \le i
< j \le t$, Theorem 2.4.3 tells us that the determinant of the matrix
is nonzero. As we also know that any nonzero element modulo $p$ (a
prime) has a multiplicative inverse, so the matrix above can be
inverted. Thus, re-constructing the polynomial $f(x)$.
\\

\textbf{$t-1$ shareholders} \\
If $t-1$ shareholders pool their shares, denoted $\{(x_1,y_1),\
(x_2,y_2),\ \ldots,\ (x_{t-1},y_{t-1})\}$, and determine that $f(0) = k'$.

\begin{equation}
  \begin{pmatrix}
    1      & x_1    & x_1^2    & \cdots & x_1^{t-1} \\
    1      & x_2    & x_2^2    & \cdots & x_2^{t-1} \\
    \cdots & \cdots & \cdots  & \cdots & \cdots \\
    1      & x_{t-1} & x_{t-1}^2 & \cdots & x_{t-1}^{t-1} \\
    1      & 0      & 0       & \cdots & 0
  \end{pmatrix}
  \begin{pmatrix}
    c_0 \\
    c_1 \\
    c_2 \\
    \cdots \\
    c_{t-1}
  \end{pmatrix} =
  \begin{pmatrix}
    y_0 \\
    y_1 \\
    y_2 \\
    \cdots \\
    k'
  \end{pmatrix} \mod p
\end{equation}

As for the same argument as above, for any value of $k'$ we can
re-construct an unique polynomial that could have generated their
shares, for which $f(0) = k'$. Hence, the shareholders have no
information about the value of the secret key.
