%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../index"
%%% End:

% Recommended:
% Why does it work?!
% Prime generation
% Modulus exponentiation
% Alice-Bob diagram

\subsection*{Agenda}
\begin{enumerate}
\item Setup
\item Proof
\item Prime generation
\item Non-determinism
\item Modular exponentiation
\end{enumerate}

\subsection{Setup}
\textbf{Public key:} $(n, e)$ where $n = pq$ for primes $p, q$ and $e \in \mathbb{Z}_{\phi(n)}^*$

\textbf{Private key:} $(p, q, d)$ where $d \in \mathbb{Z}_{\phi(n)}^*$, such that
\[ ed \equiv 1 \mod \phi(n) \]

We use $\phi(n)$ because it is hard to compute given only a large $n$, but easy with known $p, q$ since $\phi(n) = (p-1)(q-1)$, thus $d$ cannot be easily computed given just $n$ (must prime factor $n$).

\textbf{Encryption:} $m$ is the message. Encryption is
\[ Enc(m) = m^e \mod n \]

\textbf{Decryption:} $c$ is the ciphertext. Decryption is as follows:
\[ Dec(c) \equiv c^d \mod n \]
\[ Dec(c) \equiv m^{ed} \mod n \]

\textbf{Properties of RSA:}

$e$ must be chosen such that it holds that $\gcd(\phi(n),e) = 1$, where
$\phi(n) = (p - 1)(q - 1)$

$n$ must be the product of two large primes $p$ and $q$.

\subsection{Proof}
\input{lib/rsa-enc-proof.tex}

\subsection{Prime generation}
\begin{description}
\item[Trial division] $n$ is prime iff for all primes $p < \sqrt{n}$,
  $p$ does not divide $n$. Inefficient.
\item[Miller-Rabin] Probabilistic method of checking whether a number
  $n$ is prime. Can be iterated to increase probability of correctness.
\end{description}

\subsection{Weaknesses of RSA}
\input{lib/rsa-enc-weakness.tex}

\subsection{Non-determinism}
As mentioned, if $c_1$ and $c_2$ are encrypted with the same key then:
\begin{equation}
  (c_1c_2)^d \equiv (c_1^d) (c_2^d) \equiv m_1m_2 \mod n
\end{equation}
This is an unfortunate result of RSA being deterministic. El-Gamal
uses a random exponent in the encryption process and therefor does not
have the same property.

If RSA is combined with a padding algorithm containing randomness,
non-determinism is achieved.

\subsection{Modular exponentiation}

Efficient way of computing $c \equiv b^e \mod m$. Utilize that the following
to equations are equivalent
\begin{align*}
  c &\equiv ab \mod m\\
  c &\equiv (a \cdot (b \mod m)) \mod
\end{align*}

Given $c, b, e$ the algorithm is
\begin{enumerate}
\item Set $c = 1$, $e' = 0$
\item Increment $e'$
\item Set $c = b \cdot c \mod m$
\item If $e' < e$ go to 2, else $c$ is the correct answer
\end{enumerate}

The example b = 4, e = 13, and m = 497 is presented again. The algorithm passes through step 3 thirteen times:

\begin{enumerate}
\item $e' = 1 $, $c = (1 * 4) \mod 497 = 4 \mod 497 = 4$.
\item $e' = 2 $, $c = (4 * 4) \mod 497 = 16 \mod 497 = 16$.
\item $e' = 3 $, $c = (16 * 4) \mod 497 = 64 \mod 497 = 64$.
\item $e' = 4 $, $c = (64 * 4) \mod 497 = 256 \mod 497 = 256$.
\item $e' = 5 $, $c = (256 * 4) \mod 497 = 1024 \mod 497 = 30$.
\item $e' = 6 $, $c = (30 * 4) \mod 497 = 120 \mod 497 = 120$.
\item $e' = 7 $, $c = (120 * 4) \mod 497 = 480 \mod 497 = 480$.
\item $e' = 8 $, $c = (480 * 4) \mod 497 = 1920 \mod 497 = 429$.
\item $e' = 9 $, $c = (429 * 4) \mod 497 = 1716 \mod 497 = 225$.
\item $e' = 10$, $ c = (225 * 4) \mod 497 = 900 \mod 497 = 403$.
\item $e' = 11$, $ c = (403 * 4) \mod 497 = 1612 \mod 497 = 121$.
\item $e' = 12$, $ c = (121 * 4) \mod 497 = 484 \mod 497 = 484$.
\item $e' = 13$, $ c = (484 * 4) \mod 497 = 1936 \mod 497 = 445$.
\end{enumerate}
